\documentclass{ximera}


\begin{document}
Print Name: $\_\_\_\_$\\
OSU name.\#: $\_\_\_\_$\\
Instructor: $\_\_\_\_$\\
Signature: $\_\_\_\_$

\section*{MATH 1075}
\section*{Midterm Exam 2}
\section*{Autumn 2019}
\section*{Instructions:}
You have 55 minutes to complete this exam.\\
When applicable, mark answer bubbles completely like this , NOT like this or ↓ .

Circle or box all other final answers.\\
A calculator may be used given the calculator policy outlined in the syllabus. Even if a calculator is used, you must show all work on each problem to receive full credit.

Write clearly and legibly to receive credit.\\
Do not round; give only exact answers (unless specified otherwise).

Problem 1. Identify the solution(s) to the equation below. Select all that apply. (5 pts)

$$5(3 m+1)(m-2)(m+7)=0$$

◯ $m=-7$\\
◯ $m=-3$\\
◯ $m=0$\\
◯ $m=2$\\
◯ $m=5$

Problem 2. Identify the restricted value(s) of the expression below. Select all that apply. (5 pts)

$$\frac{7 t(t-4)}{2(t-1)}$$

◯ $t=0$\\
◯ $t=1$\\
◯ $t=2$\\
◯ $t=4$\\
◯ $t=7$

Problem 3. The time it takes to drive from City A to City B varies inversely with the speed of the vehicle.\\
(a) Write an equation that expresses the relationship between time, speed, and the constant of variation $(k)$.\\
(b) It takes 12 hours for a car to complete the trip driving 50 miles per hour. Find the value of $k$.\\
(c) How long would the trip take in a car driving 65 miles per hour? ( 3 pts)

\section*{Exam 2 - Form A}
Problem 4. Add. Simplify your final answer.\\
(13 pts)

$$\frac{9}{(4 x+9)(x-3)}+\frac{3}{(4 x+9)(x+4)}$$

Problem 5. Simplify the expression below.

$$\frac{\frac{14}{x-8}-7}{\frac{4}{x-8}-2}$$

Problem 6. Simplify the rational expression.

$$\frac{28 w^{4}(w+6)^{2}(3 w+2)}{32 w^{3}(w+6)^{3}}$$

Problem 7. Solve the equation using the Zero Product Rule.

$$(3 x-1)(x+4)=30$$

Problem 8. Divide. Simplify your final answer.

$$\frac{3 y}{2 a} \div \frac{12 y}{6 a y}$$

Problem 9. A certain bathtub has two faucets, one for hot water and one for cold water. If both are turned on simultaneously, the two faucets can fill the tub in 8 minutes. The hot water faucet can fill the tub by itself in 12 minutes. How long would it take the cold water faucet to fill the tub by itself?\\
(9 pts)

Problem 10. Multiply. Simplify your final answer.\\
(9 pts)\\
$\frac{x^{2}+5 x+6}{x^{2}-4 x+3} \cdot \frac{x-3}{4 x+12}$

\section*{Exam 2 - Form A}
Problem 11. Solve the equation.\\
(11 pts)

$$\frac{G+5}{G+4}-1=\frac{G-1}{G-4}$$

Problem 12. Madelyn typically walks 8 miles per hour slower than she runs. One morning, she used a stopwatch on her daily run and found that she walks 6 miles in the same amount of time that she runs 18 miles. How fast does Madelyn walk?\\
(7 pts)


\end{document}