\documentclass{ximera}

\begin{document}
\section*{Print Name:}
$\_\_\_\_$\\
BuckID Number: $\_\_\_\_$\\
Instructor: $\_\_\_\_$\\
Signature: $\_\_\_\_$

\section*{MATH 1075}
\section*{Final Exam}
\section*{Autumn 2019}
\section*{Read Instructions Carefully.}
\begin{itemize}
  \item Write clearly and legibly to receive full credit.
  \item You have 1 hour 45 minutes to complete this exam. Be careful and check your work.
  \item The exam will be scored out of 150 points, but there are 153 points possible. Extra credit will be awarded for any scores above 150.
  \item When applicable, mark answer bubbles completely, like this: Problems with answer bubbles have only ONE correct answer.
  \item Some problems require you to write your final answer on a designated line. Circle or box all other final answers.
  \item A calculator may be used given the calculator policy outlined in the syllabus. Even if a calculator is used, you must show all work on each problem to receive full credit.
\end{itemize}

Problem 1. Solve the compound inequality. Select the correct solution set from the options below.\\
(5 pts)

$$2 x-4 \leq-16 \quad \text { or } \quad 3 x+2 \geq 11$$

◯ $[6, \infty)$\\
◯ $[-6,3]$\\
◯ $(-\infty,-10] \cup[3, \infty)$\\
◯ $(-\infty,-6] \cup[3, \infty)$\\
◯ $(-\infty, \infty)$\\
◯ $\varnothing$\\
Problem 2. Solve the absolute value inequality. Select the correct solution set from the options below.\\
( 5 pts )

$$6|y+7|-8<4$$

◯ $(-9,-5)$\\
◯ $(-\infty,-5)$\\
◯ $\left(-\infty,-\frac{5}{6}\right)$\\
◯ $\left(-\infty,-\frac{22}{3}\right)$\\
◯ $(-\infty, \infty)$\\
◯ $\varnothing$\\
Problem 3. Subtract the rational expressions and simplify. Select the correct answer from the options below.\\
(5 pts)

$$\frac{x}{5 x-8}-\frac{x+4}{6 x}$$

◯ $-\frac{4}{6 x(5 x-8)}$\\
◯ $-\frac{4}{x+8}$\\
◯ $\frac{4}{x+8}$\\
◯ $\frac{(x-4)(x-8)}{6 x(5 x-8)}$\\
◯ $\frac{(x+4)(x-8)}{6 x(5 x-8)}$

Problem 4. Multiply the rational expressions and simplify. Select the correct answer from the options below.\\
(5 pts)

$$\frac{3 x}{2 a} \cdot \frac{8 a^{3} x^{5}}{15 x}$$

◯ $\frac{45}{16 a^{4} x^{3}}$\\
◯ $\frac{16 a^{4} x^{3}}{45}$\\
◯ $\frac{24 a^{3}}{x^{6}}$\\
◯ $\frac{12 a^{2} x^{5}}{15}$\\
◯ $\frac{4 a^{2} x^{5}}{5}$

Problem 5. Select the least common denominator (LCD) of the two rational expressions from among the options below.

$$\frac{14}{3 s t(s+3)^{2}(t-1)} \quad \text { and } \quad \frac{21}{9 t^{2}(s+3)^{3}(t-2)}$$

◯ $3 t(s+3)^{2}$\\
◯ $9 s t^{2}(s+3)^{3}(t-1)(t-2)$\\
◯ $9 s t^{3}(s+3)^{5}(t-1)(t-2)$\\
◯ $12 s t^{3}(s+3)^{5}(t-1)(t-2)$\\
◯ $27 s t^{3}(s+3)^{5}(t-1)(t-2)$

Problem 6. Simplify the radical expression. Select the correct answer from the options below.\\
(5 pts)

$$\sqrt[3]{108 x^{7}}$$

◯ $3 x^{2} \sqrt[3]{4 x}$\\
◯ $9 x^{2} \sqrt[3]{4 x}$\\
◯ $36 x^{2} \sqrt[3]{x}$\\
◯ $x^{2}+\sqrt[3]{4 x}+3$\\
◯ $x^{2}+\sqrt[3]{4 x}+9$

Problem 7. Simplify the sum of radical expressions. Select the correct answer from the options below.\\
(5 pts)

$$5 w \sqrt{27 u^{3}}+u \sqrt{75 u w^{2}}$$

◯ $(5 w+u) \sqrt{27 u^{3}+75 u w^{2}}$\\
◯ $5 u w \sqrt{102 u^{4} w^{2}}$\\
◯ $5 u^{3} w w^{2} \sqrt{102}$\\
◯ $20 u w \sqrt{3 u}$\\
◯ $20 u w \sqrt{6 u}$

Problem 8. Let $f(x)=\sqrt{5 x+1}$ and $g(x)=2|x-1|$. Use these functions to find the following values.\\
(3 pts each)

$$\begin{array}{ll}
(f-g)(3)=\quad & (f \circ g)(-2)=\square \\
& (f g)(3)= \\
(g \div f)(0) & (g \circ f)(7)=
\end{array}$$

Problem 9. Factor completely.\\
(6 pts)\\
$9 w^{2}-3 y-27 w+w y$

Write your answer here: $\_\_\_\_$

Problem 10. Factor completely.\\
(6 pts)\\
$y^{2}+8 y+12$

Write your answer here: $\_\_\_\_$

Problem 11. Factor completely.\\
(6 pts)\\
$3 z^{2}+14 z-24$

Write your answer here: $\_\_\_\_$

Problem 12. Factor completely.\\
(6 pts)\\
$27 x^{3}-192 x$

Write your answer here: $\_\_\_\_$

Problem 13. Solve the quadratic equation. Show your work.

$$5 x^{2}+9 x+2=0$$

Problem 14. Solve the quadratic equation. Show your work.

$$y^{2}+18 y+81=20$$

Problem 15. Solve the quadratic equation. Show your work.

$$z^{2}+3 z+2=0$$

Problem 16. When a constant force acts upon an object, the acceleration (a) of the object varies inversely with its mass ( $M$ ).\\
(a) Which equation below models the variation described above? The letter $k$ represents the constant of variation.\\
◯ $a=k M$\\
◯ $M=a k$\\
◯ $a=\frac{M}{k}$\\
◯ $a=\frac{k}{M}$\\
◯ $M=\frac{k}{a^{2}}$\\
(b) In a game of billiards, a player struck a mass of $M=156$ grams. The ball accelerated at a rate of $a=42 \mathrm{~m} / \mathrm{s}^{2}$. Find the constant of variation, $k$. (3 pts)

Write your answer here: $k=$ $\_\_\_\_$\\
(c) Suppose the billiards player were to strike a marble with mass $M=7$ grams with the same amount of force. What would be the marble's acceleration?

Write your answer here: $a=$ $\_\_\_\_$

Problem 17. Solve the rational equation.

$$\frac{2}{w+4}=\frac{8}{w+4}+6$$

Problem 18. Solve the radical equation.

$$u=\sqrt{11 u-24}$$

Problem 19. Let $f(x)=4 x^{2}+16 x+19$.\\
(a) Select the equation that represents $f(x)$ written in vertex form. (5 pts)\\
$f(x)=4 x(4 x+4)+3$\\
$f(x)=4(x-2)^{2}+3$\\
$f(x)=4(x+2)^{2}+3$\\
$f(x)=4(x+4)^{2}+19$\\
$f(x)=4(x-4)^{2}+19$\\
(b) Select the equation of the axis of symmetry of $f(x)$.\\
◯ $x=2$\\
◯ $x=-2$\\
◯ $y=2$\\
◯ $y=4$\\
◯ None of the above\\
(c) What is the minimum value of the function?

Write your answer here: $\_\_\_\_$\\
(d) At which value of $x$ does $f(x)$ reach its minimum value?

Write your answer here: $\_\_\_\_$\\
(e) Fill in the blanks with the coordinates of the vertex of $f(x)$.\\
$($ $\_\_\_\_$ , $\_\_\_\_$ )\\
(f) True or False? If $f(x)$ were graphed, the vertex of the function would lie to the left of the $y$-axis.\\
(3 pts)

\section*{TRUE FALSE}
(g) True or False? If $f(x)$ were graphed, the vertex of the function would lie above the $x$-axis.\\
(3 pts)\\
TRUE FALSE

This page was intentionally left blank.


\end{document}